# Regressions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
```

*Case Study: Iowa Corn Yields*

## Weather Data Analysis

### Load PRISM daily maximum temperatures

```{r tmax data, warning=F,message=F}

# daily max temperature
# dimensions: counties x days x years
prism <- readMat("Data/Chapter_5/prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```
```{r tidying up,warning=F}

# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)

```

## Temperature trends

### Summer temperature trends: Winneshiek County

```{r temp trends, warning=F,message=F}

#convert to numeric
tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

#filter by summer
winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

#plot data
ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

#linear model
lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)

```

### Winter Temperatures - Winneshiek County

```{r winter temps,warning=F,message=F,results='hide'}

#filter by winter
winnewinter <- tmaxdf %>%
  filter(countyfp==191 & (doy <= 59 | doy >= 335) & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

#plot data
ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

#linear model
lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)

```

## Multiple regression -- Quadratic time trend

```{r quadratic temp trend,warning=F,message=F}

#create quadratic value
winnewinter$yearsq <- winnewinter$year^2

#quadratic model
lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

#plot data
ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)")

```

## Download NASS corn yield data

```{r yield download,warning=F,message=F,results='hide'}

# set our API key with NASS
nassqs_auth(key = "5B47588E-E354-32FE-A42C-2789B2935796 ")

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```


##  Fit and Plot Linear time trend - Winneshiek County, IA

```{r,warning=F,message=F}

#filter data to Winneshiek county
winneyields <- cornyields %>%
  filter(county_name == 'WINNESHIEK')

#plot data
ggplot(winneyields, mapping = aes(x = year, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

#linear model
lm_winnieyields <- lm(yield~ year, winneyields)
summary(lm_winnieyields)
```



The adjusted R-squared of 0.7488 shows there is a significant, but not ironclad, linear relationship between year and yield.

## Fit and Plot quadratic time trend - Winneshiek County, IA

```{r,warning=F,message=F}

#create quadratic valye
winneyields$yearsq_yield <- (winneyields$year)^2

#fit quadratic model
lm_winneyields_quad <- lm(yield ~ year + yearsq_yield, winneyields)
summary(lm_winneyields_quad)
winneyields$fitted <- lm_winneyields_quad$fitted.values

#plot model
ggplot(winneyields) +
  geom_point(mapping = aes(x = year, y = yield)) +
  geom_line(mapping = aes(x = year, y = winneyields$fitted)) +
  theme_bw() +
  labs(x = "year", y = "yield")
```


The adjusted R-squared of 0.743 shows there is a significant, but not ironclad, quadratic relationship between year and yield.


## Summer Temperature and yields for the Winneshiek County time series.

```{r,warning=F,message=F}
summer_yields <- left_join(winnesummer, winneyields)

#Linear Trend
ggplot(summer_yields, mapping = aes(x = meantmax, y = yield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Tmax (°C)", y = "Yield") +
  geom_smooth(method = lm)

lm_summeryield <- lm(yield~ meantmax, summer_yields)
summary(lm_summeryield)

#Quadratic Trend

summer_yields$tempsq_summer <-(summer_yields$meantmax)^2

lm_summeryields_quad <- lm(yield ~ meantmax + tempsq_summer, summer_yields)
summary(lm_summeryields_quad)
summer_yields$fitted <- lm_summeryields_quad$fitted.values

ggplot(summer_yields) +
  geom_point(mapping = aes(x = meantmax, y = yield)) +
  geom_line(mapping = aes(x = meantmax, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "T_max")

```

The linear adjusted R-squared of 0.0041 indicates there is not a significant linear relationship between yield and average summer temperature. However, using a quadratic regression with T_max^2 increases the Adjusted r-squared to 0.1984, indicating a (relatively) more significant relationship between yield and average summer temperature.


## 2018 Cross-Section relationship between temperature and yield -- Winneshiek County

```{r,warning=F,message=F}

#filter data

yields_18 <- cornyields %>%
  filter(year==2018) %>%
  group_by(county_name) %>%
  unique() %>%
  filter(!is.na(county_ansi))

summer_county_18 <- tmaxdf %>%
  group_by(countyfp) %>%
  filter(year==2018) %>%
  filter(doy >= 152 & doy <=243) %>%
  summarize(meatmax=mean(tmax)) %>%
  rename(county_ansi = "countyfp")

summer_county_18$county_ansi <- as.numeric(as.character(summer_county_18$county_ansi))

county_summer_yield <- left_join(yields_18,summer_county_18, by='county_ansi')
```

```{r,warning=F,message=F}
#fitted values

county_summer_yield$tempsq_county_summer <-(county_summer_yield$meatmax)^2

lm_countysummeryields_quad <- lm(yield ~ meatmax + tempsq_county_summer, county_summer_yield)
summary(lm_summeryields_quad)
county_summer_yield$fitted <- lm_countysummeryields_quad$fitted.values
```

```{r,warning=F,message=F}
#construct plot
ggplot(county_summer_yield) +
  geom_point(mapping = aes(x = meatmax, y = yield)) +
  geom_line(mapping = aes(x = meatmax, y = fitted)) +
  theme_bw() +
  labs(x = "Tmax (°C)", y = "Corn Yield")

```

The adjusted R-squared of 0.1984 shows that a quadratic regression isn't terrible, but doesn't quite fit. Overall, both the regression and overall plot indicates that the highest yields take place around a t_max of 28.


## Panel Graphs for All Iowa Counties
  
```{r, warning=F,message=F}

#group by counties
county_yield_panel <- cornyieldsall %>% 
  group_by(county_name) %>% 
  unique() %>% 
  filter(!is.na(county_ansi))

#filter by summer
county_summer_panel <- tmaxdf %>%
  group_by(countyfp) %>%
  filter(doy >= 152 & doy <= 243) %>%
  summarize(meantmax = mean(tmax)) %>% 
  rename(county_ansi = "countyfp")

#creating panel dataset
county_summer_panel$county_ansi <- as.numeric(as.character(county_summer_panel$county_ansi))
county_summeryield_panel <- left_join(county_yield_panel,county_summer_panel, by='county_ansi') 
county_summeryield_panel <- subset(county_summeryield_panel, select = c(county_ansi, yield, meantmax, year))
county_summeryield_panel$county_ansi <- as.factor(county_summeryield_panel$county_ansi)
county_summeryield_panel$meantmaxsq <- county_summeryield_panel$meantmax^2
lm_county_summeryield_panel <- lm(yield ~ county_ansi + meantmax + meantmaxsq + year, county_summeryield_panel)
summary(lm_county_summeryield_panel)
county_summeryield_panel$fitted <- lm_county_summeryield_panel$fitted.values

#graphing
ggplot(county_summeryield_panel) +
  geom_point(mapping = aes(x = yield, y = fitted)) +
  geom_smooth(mapping = aes(x = yield, y = fitted),method = lm) +
  labs(x = "Actual Yield", y = "Fitted Yield")
```


The fitted yield matches the actual yield; the adjusted r-squared of 0.65 and extremely low p-value indicate the yield is a decent fit. The significance of both temperature values to the fitted model has decreased, with N/A coefficients.

### Cross-Section Plot for Soybeans -- All Iowa Counties


```{r, warning=FALSE,message=FALSE,results='hide'}
#get soybeans data
params_soy <- list(commodity_desc = "Soybeans", statisticcat_desc="YIELD", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

#create documents
soyyieldsall <- nassqs_yields(params_soy)

#setting values as numerics
soyyieldsall$county_ansi <- as.numeric(soyyieldsall$county_ansi)
soyyieldsall$yield <- as.numeric(soyyieldsall$Value)

#filter relevant values
soyyields <- select(soyyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
soyyields <- tibble(soyyields)

#cross-section
soy_yields_18 <- soyyields %>%
  filter(year==2018) %>%
  group_by(county_name) %>%
  unique() %>%
  filter(!is.na(county_ansi))

summer_county_18 <- tmaxdf %>%
  group_by(countyfp) %>%
  filter(year==2018) %>%
  filter(doy >= 152 & doy <=243) %>%
  summarize(meatmax=mean(tmax)) %>%
  rename(county_ansi = "countyfp")

summer_county_18$county_ansi <- as.numeric(as.character(summer_county_18$county_ansi))

#join datasets
soy_county_summer_yield <- left_join(soy_yields_18,summer_county_18, by='county_ansi')

soy_county_summer_yield$tempsq_county_summer <-(soy_county_summer_yield$meatmax)^2

```
```{r, message=F,warning=F}

#fit linear model
lm_soy_countysummeryields_quad <- lm(yield ~ meatmax + tempsq_county_summer, soy_county_summer_yield)
summary(lm_summeryields_quad)
soy_county_summer_yield$fitted <- lm_soy_countysummeryields_quad$fitted.values
```
```{r}
#construct plot
ggplot(soy_county_summer_yield) +
  geom_point(mapping = aes(x = meatmax, y = yield)) +
  geom_line(mapping = aes(x = meatmax, y = fitted)) +
  theme_bw() +
  labs(x = "Tmax (°C)", y = "Soy Yield")
```


A county cross-section for soil yield and temperature shows a similar relationship to corn yield. The highest yields are around a T-max of 28, and can be reasonably modeled by a quadratic regression.
